---
title: 'Data 624 - Homework #1'
author: "Paul Britton"
date: ''
output:
  html_document:
    df_print: paged
---


Please submit exercises 2.1, 2.2, 2.3 and 2.6 from the Hyndman online Forecasting book.  Please submit both your Rpubs link as well as attach the .rmd file with your code.
 
 
```{r}
rm(list = ls())
library(forecast)
library(readxl)
library(RCurl)
library(fpp2)

```

## Question 2.1

```{r}
describe.data <- function(data) { 
  freq <- frequency(data)
  outlier <- which.max(data)
  return(c(freq,outlier))
}

```

### Gold

```{r}
help(gold)
autoplot(gold)
question1 <- describe.data(gold)
```

- The "gold" data represents the daily (morning) gold price in USD for the period 1985-01-01 to 1989-03-31.
- The frequency of this data is `r question1[1]`.  The data is daily.
- The outlier datapoint appears at index `r question1[2]` and corresponds to a price of `r gold[question1[2]]`


### Wollyrnq

```{r}
help(woolyrnq)
autoplot(woolyrnq)
question2 <- describe.data(woolyrnq)
```


- The "woolyrnq" data represents the quarterly production of woolen yard (tonnes) in Australia for the period Mar-1965 to Sep-1994
- The frequency of this data is `r question2[1]` observation per year, i.e. quarterly.

### Wollyrnq

```{r}
help(gas)
autoplot(gas)
question3 <- describe.data(gas)
```

- The "gas" dataset shows Australian monthly gas production (units not specified) for the period 1956-1995 
- The frequency of this data is `r question2[1]` observation per year, i.e. monthly.


## Question 2.2

### A

```{r}

tute1 <- read.csv("http://otexts.com/fpp2/extrafiles/tute1.csv",header=T)
View(tute1)

```

### B

```{r}

mytimeseries <- ts(tute1[,-1], start=1981, frequency=4)
```

### C
```{r}

autoplot(mytimeseries, facets=TRUE,main="With 'Facets' Argument")

autoplot(mytimeseries,main="Without 'Facets' Argument")


```


## Question 2.3

```{r}

retaildata <- readxl::read_excel("C:/Users/Paul/OneDrive - CUNY School of Professional Studies/CUNY/DATA 624/retail.xlsx", skip=1)

my.ts <- ts(retaildata[,"A3349388W"],
  frequency=12, start=c(1982,4))

autoplot(my.ts)
ggseasonplot(my.ts)
ggsubseriesplot(my.ts)
gglagplot(my.ts)
ggAcf(my.ts)


```

For this question I just picked a column at random and ended up with "Turnover ;  Total (State) ;  Takeaway food services ;".

* As per the timeseries plot, there is a clear upward sloping trend in the data and there are definately some seasonal / cyclical effects apparent
* The seasonaly plot shows a few things:  a consistent dip in Feb, a rise towards the end of of the year. The same is present in the sub-season plot.
* The lag plots are difficult to look at "as is" because the seasonal effects are likely weak.  I wonder whether it would make more sense to difference the timeseries before applying this plotting function
* The ACF plot shows a high degree of auto-correlation with low decay.



## Question 2.6

Use the following graphics functions: autoplot(), ggseasonplot(), ggsubseriesplot(), gglagplot(), ggAcf() and explore features from the following time series: hsales, usdeaths, bricksq, sunspotarea, gasoline.

Can you spot any seasonality, cyclicity and trend?
What do you learn about the series?

### hsales


```{r}
my.ts <- hsales

autoplot(my.ts)
ggseasonplot(my.ts)
ggsubseriesplot(my.ts)
gglagplot(my.ts)
ggAcf(my.ts)

```

* As per the timeseries plot, there is no obvious trend, but it appears as though there may be some seasonality, and there absolutely appears to be some cyclicality - large long-term osscilations between 30 and 80.
* The seasonality plots show a an uppward effect between Jan and March followed by a downward slope for the remainder of the year.  This is more apparent in the seasonality plot - the subseries plots don't work too well for this data.
* The ACF plot show a high degree of autocorrelation for short lookbacks (1-2 periods) and appear to capture some of the seasonal pattern thereafter.

### usdeaths


```{r}
my.ts <- usdeaths

autoplot(my.ts)
ggseasonplot(my.ts)
ggsubseriesplot(my.ts)
gglagplot(my.ts)
ggAcf(my.ts)

```
* Once again, these data are clearly seasonal and show little trend.
* The seasonality plots shows a dip in Feb, then a clear rise to a peak in Jul.  Thereafter, it drops off slightly and seems to flatten our towards the end of the year.
* The lag plots are informative here also with Feb consistently appearing at the bottim and Jul, near the top.
* The ACF almost looks like a sine wave, indicative of a pattern in the data.




### bricksq


```{r}
my.ts <- bricksq

autoplot(my.ts)
ggseasonplot(my.ts)
ggsubseriesplot(my.ts)
gglagplot(my.ts)
ggAcf(my.ts)

```

* This data shows an upward trend and a seemingly regular moves around that trend.
* The seasonality plots seem to indicate that Q1 is the lowest point in there year, otherwise the data appear to be somewhat flat.
* The ACF shows reasonably strong relationships with slow decay



### sunspotarea


```{r}
my.ts <-  sunspotarea

autoplot(my.ts)
#ggseasonplot(my.ts)
#ggsubseriesplot(my.ts)
gglagplot(my.ts)
ggAcf(my.ts)

```

* This data shows a clear cycle and no trend.
* Given that the data is annual observations, there is no "seasonal" effect, however the cycle, which appears to be about 10 years, is quite apparent.
* The ACF is also indicative of a cyclical pattern in that it looks like a sine wave with nadirs and peaks every 5 & 10 years, respectively.



### gasoline


```{r}
my.ts <-  gasoline

autoplot(my.ts)
ggseasonplot(my.ts)
#ggsubseriesplot(my.ts)
gglagplot(my.ts)
ggAcf(my.ts)

```

* This data shows a noisy up trend with apparent seasonality.
* The seasonal plot shows a peak near weeks 0-4 and a lull for the few weeks following.  It's doesn't appear terribly strong.
* The ACF shows a slow and cosistent decay and has an odd pattern... possibly of an annual season.
* My gut feel is that this data might be easier to digest as Monthly or Quarterly - 
